---
title: "eds221-day8-comp1"
format: html
editor: visual
---

# Review Basics of ggplot

```{r}
library(tidyverse)
library(here)
library(janitor)

library(patchwork)
library(ggrepel)
library(gghighlight)
library(paletteer)
```

# Read in data
```{r}
lizards <- read_csv(here::here("data_tidy", "lizards.csv"))
```

data wrangling to create a new subset from izards where you:
1. limit the data to only contain western whiptails, colorado checkered, and eastern fence lizards
2. add a new column with total length converted to cm
3. group by species and site
4. return a summary table with mean total length and standard deviation of total length by species and site
```{r}
lizards_limit <- lizards %>% 
  filter(common_name %in% c("western whiptails", "colorado checkered", "eastern fence")) %>% 
  mutate(length_cm = total_length / 10) %>% 
  group_by(common_name, site) %>% 
  summarise(mean_total_length = mean(total_length),
            sd_total_length = sd(total_length))
```

# 

```{r}
ggplot(data = lizards, aes(x = total_length, y = weight)) + 
  geom_point(aes(color = weight)) + 
  scale_y_continuous(limits = c(0, 100), 
                     breaks = seq(from = 0, to = 100, by = 20),
                     expand = c(0,0)) + 
  scale_color_gradient(low = "pink", high = "purple")
  
#  facet_wrap(~ common_name, scales = "free")
```

```{r}
lizard_counts <- lizards %>% 
  dplyr::count(common_name)

#or use group_by(common_name), summarize(n = n()). This helps for means, sd, and easier to name columns
#use dplyr:: count to find the total number of lizards by common name in the lizards dataset
```


```{r}
ggplot(data = lizard_counts, aes(x = n, y = fct_reorder(common_name, n))) + 
  geom_col()
```





